---
title: Counting transaction items
author: Carter Rees
date: '2017-11-11'
slug: counting-items-in-your-transaction-basket
categories:
  - R
tags:
  - dplyr
  - tidyr
---

In this post I will show you how to count the number of items a user has selected (or bought, or triggered) within a data feed. 

Transaction data can be a treasure trove of user behavior if you can get it into the right format for counting. I have put together a simple example below on how you can use R to get counts of user click behaviors at various levels of aggregation. Note: the data are entered in a sorted format. You could easily accomplish this by using the arrange command within dplyr.

```{r}
# Setup example data
customer_id <- c("aid123","aid123","aid123", "aid123",
                 "aid432","aid432",
                 "aid887","aid887","aid887","aid887","aid887",
                 "aid98","aid98", "aid98")
visit_num <- c("1", "1", "1", "2",
               "1", "1",
               "1", "2", "2", "3", "4",
               "1", "1", "1")
item <- c("socks", "socks", "shoes", "shoes",
          "tires", "brakes", 
          "shoes","shoes", "shoes", "shoes", "shoes",
          "trimmer", "trimmer", "trimmer")
transactions <- data.frame(customer_id,visit_num, item)
```
Now that the data is set up, we can count items by visit number by visitor id and then a higher level of aggregation at just visitor id. You can imagine that if you had a date value the aggregation could by visitor id by date of which the result would be daily user behavior counts.

```{r message = FALSE}
library(dplyr)
library(tidyr)
```

```{r}
# Count items by visitor id and visit number.
purchases_v1 <- transactions %>% group_by(customer_id, visit_num) %>% 
  count(item) %>%
  spread(item, n, fill=0)

print(purchases_v1)
```
The table above accurately reflects the transaction data. Visitor aid123 bought 1 pair of shoes and 2 pair of socks on the first visit.

The follwing code will create a summary of purchases at the vistor level regardless of visit number.

```{r}
# Count items by visitor id and visit number.
purchases_v2 <- transactions %>% group_by(customer_id) %>% 
  count(item) %>%
  spread(item, n, fill=0)

print(purchases_v2)

```

We can now see that visitor aid887 had 5 shoe purchases that were spread over 4 visits but are now aggregated to reflect overall purchase behavior. 


